Stuff

### 2014-01-04 paper/System.tex
added more explanations

Each cache has a \emph{coherence state} associated with each address. The
coherence state can be one of the three values: \Mo (Modified), \Sh (Shared) or
\In (Invalid). \Mo state denotes that the subtree rooted at the cache in \Mo
state for that address can modify the data, \ie both read and write the data
corresponding to the address. \Sh state denotes that the subtree rooted at the
cache in \Sh state for that address can only read the data corresponding to
that address and \In state denotes that the subtree rooted at the cache in \In
state can neither read nor write the data corresponding to the address. The
meanings of the state is clear, but one must verify the actual cache coherence
protocol for violations of these invariants, which is what the proof in this
paper intends to do. There's a logical ordering between the states, in terms of
the permissions that they have: $\Mo > \Sh > \In$ (the notation should be read
assuming that the order is transitively closed).
In addition, each non-leaf cache maintains a \emph{directory}, which contains a
local version of the state of each of its children for a particular address.
This local version may be stale, but it should not be lower than the real state
of the child (again this property has to be proven for a cache coherence
protocol)

@@

Figure \ref{format} shows the message format transmitted in the 3 channels
between caches. The messages transmitted between the processor and the L1 cache
is the same as in Figure \ref{req-resp}.
\begin{figure}
\centering
\begin{tabular}{|l|lp{.9\textwidth}|}
\hline
\multirow{2}{*}{\cpReq} & \from: & Current state of child\\
& \myto: & Desired (higher) state of child\\
\hline
\multirow{3}{*}{\cpResp} & \from: & Current state of child\\
& \myto: & Downgraded state of child\\
& \data: & Data associated with the address, if necessary\\
\multirow{4}{*}{\pc} & \type: & \Req{} (for requests) and \Resp{} (for responses)\\
\hline
& \from: & Current state of parent's directory for the child\\
& \myto: & Desired (lower) directory state of a child for \Req{} and upgraded
directory state of a child for \Resp{}\\
& \data: & Data associated with the address if necessary (for \Resp{} only)\\
\hline
\end{tabular}
\caption{Data types messages between caches}
\label{format}
\end{figure}

>>>>

Each cache has a \emph{coherence state} associated with each address. The
coherence state can be one of the three values: \Mo{} (Modified), \Sh{} (Shared) or
\In{} (Invalid). \Mo{} state denotes that the subtree rooted at the cache in \Mo{}
state for that address can modify the data, \ie{} both read and write the data
corresponding to the address. \Sh{} state denotes that the subtree rooted at the
cache in \Sh{} state for that address can only read the data corresponding to
that address and \In{} state denotes that the subtree rooted at the cache in \In{}
state can neither read nor write the data corresponding to the address. The
meanings of the state is clear, but one must verify the actual cache coherence
protocol for violations of these invariants, which is what the proof in this
paper intends to do. There's a logical ordering between the states, in terms of
the permissions that they have: $\Mo > \Sh > \In$ (the notation should be read
assuming that the order is transitively closed).
In addition, each non-leaf cache maintains a \emph{directory}, which contains a
local version of the state of each of its children for a particular address.
This local version may be stale, but it should not be lower than the real state
of the child (again this property has to be proven for a cache coherence
protocol).

Though this paper restricts the formal discussion to the simple MSI protocol,
extension of this proof technique as well as major portions of the proof to
other protocols like MOSI and MOESI is straightforward.

@@

Figure \ref{format} shows the message format transmitted in the 3 channels
between caches. The messages transmitted between the processor and the L1 cache
is the same as in Figure \ref{req-resp} (except for the omission of the \proc{}
field, which is indicated by the channel's index).
\begin{figure}
\centering
\begin{tabular}{|l|lp{.9\textwidth}|}
\hline
\multirow{3}{*}{\cpReq} & \from: & Current state of child\\
& \myto: & Desired (higher) state of child\\
& \addr: & Address associated with the request\\
\hline
\multirow{4}{*}{\cpResp} & \from: & Current state of child\\
& \myto: & Downgraded state of child\\
& \addr: & Address associated with the request\\
& \data: & Data associated with the address, if necessary\\
\multirow{4}{*}{\pc} & \type: & \Req{} (for requests) and \Resp{} (for responses)\\
\hline
%& \from: & Current state of parent's directory for the child\\
& \myto: & Desired (lower) directory state of a child for \Req{} and upgraded
directory state of a child for \Resp{}\\
& \addr: & Address associated with the request\\
& \data: & Data associated with the address if necessary (for \Resp{} only)\\
\hline
\end{tabular}
\caption{Data types messages between caches}
\label{format}
\end{figure}

The description of the behavior of each of the caches can be given in terms of
``rules'' for state transitions. Each rule specifies a series of actions that
are performed atomically, and during the execution of a rule, no other
(conflicting) rule interferes with it. Thus, the behavior of any rule-based
system can be understood as if the rules executed one at a time.

Such a rule-based description can be automatically converted into synchronous
hardware, for instance using Bluespec \cite{Hoe:TCAD,HoeArvind:TRSSynthesis1}.
Bluespec executes each atomic rule into a single hardware cycle, which
simultaneously executing non-conflicting rules \ie{} rules that do not access the
same state (the exact conditions for simultaneous scheduling are specified in
the citations). Though the rules are logically executed atomically, their execution
can be spread across multiple cycles (as long as other conflicting rules are
not executed during the whole period of current rule's execution). Synchronous
hardware generation for multi-cycle atomic rules has also been studied
\cite{Karczmarek}.
